---
title: "Report Analysis: mpg - automatic - manual"
author: "Dom Fernandez - Motr Trend, Sr. Analyst"
date: "Saturday, June 21, 2014"
output: html_document
---

Executive Summary
========================================================
In this Report your Senior Analyst analyses two questions for you:

- Is an automatic or manual-transmission car better, when considering `MPG` - consumption?
- What is - `quantifying` - the difference in `MPG` between automatic and manual transmissions?

Using data available in the `mtcars` dataset and an initial exploratory analysis, we see that there is a big difference between the means of miles/gallon - comparing automatic-transmission v/s manual cars.

<h5> Resuts indicate: automatic-transmission = `17.15` - manual = `24.39`</h5>
--------------------------------------------------------

Analysis
--------------------------------------------------------
We start by loading the **mtcars** data frame and recode **am** to a logical **automatic** variable.

```{r echo=TRUE, eval=TRUE}
data( mtcars )
mtcars$automatic <- mtcars$am == 0
attach( mtcars )
```

Ordinary Least Square `(OLS)` model is estimated by analysing the distribution of the `automatic`-variable.

```{r echo=TRUE, eval=TRUE}
ols <- lm( mpg ~ automatic, data=mtcars )
summary( ols )
```

There is a significant negative effect of automatic on `mpg`.

Redoing the estimation with Two-Stage Least Squares (`TSLS`), using weight (**wt**) as an instrument for automatic.

```{r echo=TRUE, eval=TRUE}
tsls.fs <- lm(automatic ~ wt)
tsls <- lm(mpg ~ tsls.fs$fitted.values)
summary(tsls)
```

This coefficient is even more negative, and even more significant.

Finally the residuals are plotted for both estimations
- please refer to `Appendix`.

The residuals of the OLS estimation appear to be normally distibuted, the TSLS somewhat less, though **n** is too small to make any real claims.

Appendix: Figures
--------------------------------------------------------
```{r echo=TRUE, eval=TRUE, fig.width=10, fig.height=3}
par( mfrow=c(1,2) )
hist( mpg[which(automatic == TRUE)] )
hist( mpg[which(automatic == FALSE)] )
```

OLS and TSLS Residual-estimations.

```{r echo=TRUE, eval=TRUE, fig.width=10, fig.height=3}
par( mfrow=c(1,3) )
hist( ols$residuals )
hist(tsls.fs$residuals)
hist( tsls$residuals )
```
